---
title: "COVID-19 Preprints - published"
output: github_document
---

# Background

This file contains code used to harvest information on COVID-19 related preprints (as colected in [Covid19 Preprints](https://github.com/nicholasmfraser/covid19_preprints) which are linked to journal publications.

Currently, data on preprint-publication links is harvested from one source only:

1. Crossref (using the [rcrossref](https://github.com/ropensci/rcrossref) package)

A description of the methods for harvesting data is provided in each relevant section below.

# Load required packages 

```{r message = FALSE, warning = FALSE}

library(lubridate)
library(rcrossref)
library(tidyverse)
library(jsonlite)
```


# Set sample date
# Retrieve the latest sample date for preprints

```{r}
sample_date <- Sys.Date()

sample_date_preprints <- fromJSON(
  "https://raw.githubusercontent.com/nicholasmfraser/covid19_preprints/master/data/metadata.json"
  ) 

sample_date_preprints <- sample_date_preprints %>%
  .$sample_date %>%
  as.Date()

```


# Import dataset on preprints

This script uses the dataset on preprints collated here: [Covid19 Preprints](https://github.com/nicholasmfraser/covid19_preprints)

```{r}

covid_preprints <- read_csv("https://raw.githubusercontent.com/nicholasmfraser/covid19_preprints/master/data/covid19_preprints.csv")

# Filter to preprints with DOI
# Select columns of interest
covid_preprints <- covid_preprints %>%
  filter(identifier_type == "DOI") %>%
  select(source,identifier, posted_date)

# Quering Crossref with all DOIs gets NULL responses for DataCite DOIs -> these can be filtered out

```

# Crossref

Harvesting of Crossref metadata is carried out using the [rcrossref](https://github.com/ropensci/rcrossref) package for R. The `cr_works_` function is used to retrieve all metadata related to Metadata were retrieved for all dois in the dataset of COVID19 preprints. Note that here, the 'low level' `cr_types_` function is used to return all metadata in list format, as this also includes the field 'relation' that is not returned by the 'high level' `cr_works` function.

To identify to the Crossref API and get access to a dedicated API cluster for improved performance, an email address is stored as environment variable in .Renviron 

```{r message = FALSE, warning = FALSE, cache = TRUE}

# Set email as variable "crossref_email" in .Renviron
# Restart R session after saving .Renviron 

# Query dois
dois <- covid_preprints %>%
  pull(identifier)

cr_dois <- cr_works_(dois,
                     parse = TRUE,
                     .progress = "time")

```

Relevant preprint metadata fields are parsed from the list format returned in the previous step, to a more manageable data frame. 

```{r message = FALSE, warning = FALSE, cache = TRUE}

# Function to parse Crossref preprint data to data frame
parseCrossrefDOIs <- function(item) {
  tibble(
    DOI = item$DOI,
    is_preprint_of = if(length(item$relation$`is-preprint-of`)) "is_preprint_of" else NA_character_,
    preprint_of_doi = if(length(item$relation$`is-preprint-of`)) item$relation$`is-preprint-of`[[1]]$id  else NA_character_)
}

# Select element 'message', remove NULL elements
# This removes NULL results from DataCite DOIs
cr_dois_message <- map(cr_dois, "message") %>%
  compact()

# Iterate over posted-content list and build data frame
cr_dois_df <- map_dfr(cr_dois_message, parseCrossrefDOIs)

rm(cr_dois, cr_dois_message)

```


The dataset containing all Covid19 related preprints and the dataset containing information on published articles (from Crossref) are merged and to get a final dataset containing all Crossref preprints with information on linked published articles

```{r message = FALSE, warning = FALSE, cache = TRUE}

covid_preprints_published <- covid_preprints %>%
  right_join(cr_dois_df, by = c("identifier" = "DOI")) %>%
  rename(preprint_of_doi = preprint_of_DOI)

covid_preprints_published %>%
  write_csv("data/covid19_preprints_published.csv")


count <- covid_preprints_published %>%
  #filter(!is.na(is_preprint_of)) %>%
  count(source)


  
```

# Create metadata file (json file with sample date and release date)

```{r message = FALSE, warning = FALSE}

# Set system date as release date
release_date <- Sys.Date()

# Create metadata as list
metadata <- list()
metadata$release_date <- release_date
metadata$sample_date <- sample_date
metadata$sample_date_preprints <- sample_date_preprints
metadata$url <- "https://github.com/bmkramer/covid19_preprints_published/blob/master/data/covid19_preprints_published.csv?raw=true"

# Save as json file
metadata_json <- toJSON(metadata, pretty = TRUE, auto_unbox = TRUE)
write(metadata_json, "data/metadata.json")

```

# Visualizations

```{r message = FALSE, warning = FALSE}
# Default theme options
theme_set(theme_minimal() +
          theme(text = element_text(size = 12),
          axis.text.x = element_text(angle = 90, vjust = 0.5),
          axis.title.x = element_text(margin = margin(20, 0, 0, 0)),
          axis.title.y = element_text(margin = margin(0, 20, 0, 0)),
          legend.key.size = unit(0.5, "cm"),
          legend.text = element_text(size = 8),
          plot.caption = element_text(size = 10, hjust = 0, color = "grey25", 
                                      margin = margin(20, 0, 0, 0))))
# Create a nice color palette
pal_1 <- colorspace::lighten(pals::tol(n = 10), amount = 0.2)
pal_2 <- colorspace::lighten(pals::tol(n = 10), amount = 0.4)
palette <- c(pal_1, pal_2)
```


```{r message = FALSE, warning = FALSE}
# Weekly preprint counts

covid_preprints_published %>%
  mutate(
    status = case_when(
      is.na(is_preprint_of) ~ "not linked to published journal article",
      !is.na(is_preprint_of) ~ "linked to published journal article"
    ),
    status = factor(status),
    posted_week = ymd(cut(posted_date,
                          breaks = "week",
                          start.on.monday = TRUE))) %>%
  count(status, posted_week) %>%
  ggplot(aes(x = posted_week, y = n, 
             fill = forcats::fct_rev(status))) +
  geom_col() +
  labs(x = "Posted Date (week beginning)", y = "Preprints", fill = "status (from Crossref)",
       title = "COVID-19 preprints per week in Crossref", 
       subtitle = paste0("(preprints up until ", sample_date_preprints, ", sample date ", sample_date,")")
       ) +
  scale_x_date(date_breaks = "1 week",
               expand = c(0, 0),
               limits = c(ymd("2020-01-13"), ymd(sample_date_preprints))) +
  scale_fill_manual(values = palette) +
  ggsave("outputs/figures/covid19_preprints_published_week.png", width = 12, height = 6)
```


```{r message = FALSE, warning = FALSE}
# Cumulative daily preprint counts
covid_preprints_published %>%
  mutate(
    status = case_when(
      is.na(is_preprint_of) ~ "not linked to published journal article",
      !is.na(is_preprint_of) ~ "linked to published journal article"
    ),
    status = factor(status)) %>%
  count(status, posted_date) %>%
  complete(posted_date, nesting(status), fill = list(n = 0)) %>%
  group_by(status) %>%
  arrange(posted_date) %>%
  mutate(cumulative_n = cumsum(n)) %>%
  ggplot() +
  geom_area(aes(x = posted_date, 
                y = cumulative_n, 
                fill = forcats::fct_rev(status))) +
  labs(x = "Posted Date (week beginning)", y = "Preprints", fill = "status (from Crossref)",
       title = "COVID-19 preprints in Crossref (cumulative)", 
       subtitle = paste0("(preprints 1up until ", sample_date_preprints, ", sample date ", sample_date, ")")) +
  scale_y_continuous(labels = scales::comma) +
  scale_x_date(date_breaks = "1 week",
               expand = c(0.01, 0),
               limits = c(ymd("2020-01-13"), ymd(sample_date_preprints) + 1)) +
  scale_fill_manual(values = palette) +
  ggsave("outputs/figures/covid19_preprints_published_day_cumulative.png", width = 12, height = 6)
```

```{r message = FALSE, warning = FALSE}
# For graphs per preprint server

# Group all OSF preprints together
OSF_names <- covid_preprints_published %>%
  count(source) %>%
  filter(str_detect(source, "OSF")) %>%
  arrange(desc(n)) %>%
  pull(source)

covid_preprints_published_viz <- covid_preprints_published %>%
  mutate(source = case_when(
    source %in% OSF_names ~ "OSF",
    TRUE ~ source))

# Create vector with sources in order of descending number of preprints linked to published articles
sources_count <- covid_preprints_published_viz %>%
  filter(!is.na(is_preprint_of)) %>%
  count(source) %>%
  arrange(desc(n)) %>%
  pull(source)

```


```{r message = FALSE, warning = FALSE}

# Weekly preprint counts - per preprint server

# Select source to display in graph
var <- sources_count[8]

# Create graph
covid_preprints_published_viz %>%
  filter(source == var) %>%
  mutate(
    status = case_when(
      is.na(is_preprint_of) ~ "not linked to published journal article",
      !is.na(is_preprint_of) ~ "linked to published journal article"
    ),
    status = factor(status),
    posted_week = ymd(cut(posted_date,
                          breaks = "week",
                          start.on.monday = TRUE))) %>%
  count(status, posted_week) %>%
  ggplot(aes(x = posted_week, y = n, 
             fill = forcats::fct_rev(status))) +
  geom_col() +
  labs(x = "Posted Date (week beginning)", y = "Preprints", fill = "status (from Crossref)",
       title = paste0("COVID-19 preprints per week on ",var), 
       subtitle = paste0("(preprints up until ", sample_date_preprints, ", sample date ", sample_date,")")
       ) +
  scale_x_date(date_breaks = "1 week",
               expand = c(0, 0),
               limits = c(ymd("2020-01-13"), ymd(sample_date_preprints))) +
  scale_fill_manual(values = palette) +
  ggsave(paste0("outputs/figures/covid19_preprints_published_",
                var,
                "_week.png"), 
         width = 12, height = 6)
```