---
title: "COVID-19 Preprints - Europe PMC"
output: github_document
---

Compare links from preprints to published papers in EuropePMC
and compare to data in Crossref (all preprints servers) and bioRxiv/medRxiv

This approach queries EuropePMC for covid-related preprints directly, 
rather than only for dois in covid_preprints, to retrieve all versions

# Load required packages 

```{r message = FALSE, warning = FALSE}

library(lubridate)
library(tidyverse)
library(europepmc)
library(colorspace)

```


# Set sample date
# Retrieve the latest sample date for preprints

```{r message = FALSE, warning = FALSE}
sample_date <- Sys.Date()
sample_date <- "2021-01-27"

sample_date_preprints <- fromJSON(
  "data/metadata.json"
  ) 

sample_date_preprints <- sample_date_preprints %>%
  .$sample_date_preprints %>%
  as.Date()

```

#query EuropePMC for COVID preprints

```{r message = FALSE, warning = FALSE, cache = TRUE}

dois <- covid_preprints_published %>%
  #filter(source == "bioRxiv",
  #       !is.na(is_preprint_of)) %>%
  pull(identifier)
 
epmc_res <- epmc_search_by_doi(doi = dois, output = "raw")

epmc_preprints_res <- epmc_search(
  #use same query as used for covid_preprints.csv
  query = '(coronavirus OR covid-19 OR sars-cov OR ncov-2019 OR 2019-ncov OR hcov-19 OR sars-2) AND SRC:PPR',
  #to replicate results from the website, synonym must be set to T
  synonym = TRUE,
  #shortcut: manually check number of results on website
  #limit cannot exceed expected number of results
  limit = 26790,
  output = "raw")

parseEuropePMC_preprints <- function(item) {
  tibble(
    epmc_doi = if(length(item$doi)) item$doi else NA_character_,
    epmc_id = if(length(item$id)) item$id else NA_character_,
    epmc_source = if(length(item$source)) item$source else NA_character_,
    epm_links = length(item$commentCorrectionList$commentCorrection),
     epm_links_id = if(length(item$commentCorrectionList$commentCorrection)) item$commentCorrectionList$commentCorrection[[1]]$id else NA_character_,
    epm_links_source = if(length(item$commentCorrectionList$commentCorrection)) item$commentCorrectionList$commentCorrection[[1]]$source else NA_character_,
    epm_links_type = if(length(item$commentCorrectionList$commentCorrection)) item$commentCorrectionList$commentCorrection[[1]]$type else NA_character_
    )
}


epmc_preprints_df <- map_df(epmc_preprints_res,
                            parseEuropePMC_preprints) 


#check for cases where 1st link != "Preprint of"
check <- epmc_preprints_df %>%
  count(epm_links_type)

write_csv(epmc_preprints_df, "covid_preprints_all_epmc.csv")
epmc_preprints_df <- read_csv("covid_preprints_all_epmc.csv")

```


#Join with Crossref and CSHL data

```{r message = FALSE, warning = FALSE}

# read files to join
cshl_df <- read_csv("../../data/covid19_preprints_published_cshl.csv")
cr_df <- read_csv("../../data/covid19_preprints_published.csv")

#prep dataframes to join
#To do: turn de-versioning into a function to reuse

epmc_join <- epmc_preprints_df %>% 
  #filter to records with doi (NB this excludes e.g. arxiv)
  filter(!is.na(epmc_doi)) %>%
  select(epmc_doi, epm_links_type) %>%
  mutate(epmc_doi = str_trim(str_to_lower(epmc_doi))) %>%
  rename(identifier = epmc_doi,
         linked_epmc = epm_links_type) %>%
  mutate(linked_epmc = case_when(
    !is.na(linked_epmc) ~ "linked_epmc",
    is.na(linked_epmc) ~ NA_character_)) %>%
  mutate(in_epmc = "in_epmc") %>%
  #strip version info for clean matching
  # Remove .v2, /v2 and -v2 at the end of a string
  # with a maximum of 3 digits
  mutate(identifier_clean = str_remove(
    identifier,
    "[\\.\\/\\-]v[0-9]{1,3}$")) %>%
  select(-identifier) %>%
  #make sure duplicate identifier_clean have some linked_epmc info
  group_by(identifier_clean) %>%
  fill(linked_epmc, .direction = "downup") %>%
  ungroup() %>%
  #remove any duplicate identifier_clean
  distinct()

cshl_join <- cshl_df %>%
  select(identifier, is_published_CSHL) %>%
  mutate(identifier = str_trim(str_to_lower(identifier))) %>%
  rename(linked_cshl = is_published_CSHL) %>%
  mutate(
    linked_cshl = case_when(
      linked_cshl == TRUE ~ "linked_cshl",
      linked_cshl == FALSE ~ NA_character_))
#NB no version numbers in cshl dois

cr_join <- cr_df %>%
  select(source, identifier, posted_date, is_preprint_of) %>%
  mutate(identifier = str_trim(str_to_lower(identifier))) %>%
  rename(linked_cr = is_preprint_of) %>%
  mutate(
    linked_cr = case_when(
      !is.na(linked_cr) ~ "linked_cr",
      is.na(linked_cr) ~ NA_character_)) %>%
  #strip version info for clean matching
  # Remove .v2, /v2 and -v2 at the end of a string
  # with a maximum of 3 digits
  mutate(identifier_clean = str_remove(
    identifier,
    "[\\.\\/\\-]v[0-9]{1,3}$")) %>%
  select(source, identifier, identifier_clean, everything())
  
  
# join together   
covid_preprints_published_epmc <- cr_join %>%
  left_join(cshl_join, by = ("identifier")) %>%
  left_join(epmc_join, by = ("identifier_clean"))

write_csv(covid_preprints_published_epmc, "covid_preprints_published_epmc.csv")

#read in with column specification for linked_cshl as that has only NAs in part that is checked for guessing
covid_preprints_published_epmc <- read_csv(
  "covid_preprints_published_epmc.csv",
  col_types = cols(.default = "?", linked_cshl = "c"))

```

# Group OSF preprint servers together

```{r}
# Group all OSF preprints together
OSF_names <- covid_preprints_published_epmc %>%
  count(source) %>%
  filter(str_detect(source, "OSF")) %>%
  arrange(desc(n)) %>%
  pull(source)

covid_preprints_published_epmc2 <- covid_preprints_published_epmc %>%
  mutate(source = case_when(
    source %in% OSF_names ~ "OSF",
    TRUE ~ source))

```


#Some calculations

```{r message = FALSE, warning = FALSE}

epmc_per_source <- covid_preprints_published_epmc %>%
  group_by(source) %>%
  summarise_all(~ sum(!is.na(.))) %>%
  arrange(desc(identifier))

epmc_per_source <- epmc_per_source %>%
  rename(doi = identifier,
         linked_in_crossref = linked_cr,
         linked_in_europepmc = linked_epmc,
         linked_in_biorxiv_medrxiv = linked_cshl)

write_csv(epmc_per_source, "epmc_per_source.csv")
  
```

``

